---
title: "Regressionr"
author: "RUPESH ANUSURI"
date: "Submission time: May 26, 2022 at 11:00PM EST"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r, load_packages}
library(tidyverse)
library(caret)
library(modeldata)
library(tidymodels)
library(pls) # for pcr analysis
library(vip) # for variable importance
library(Metrics)
library(glmnet)
library(xgboost)
library(reshape2)
library(patchwork)
library(yardstick)
```

### Getting the dataset:

```{r}
data(meats)

#Removing other dependent variables
meats <- subset (meats, select = -c(fat,protein))
```

### Splitting into test data and train data with stratified sampling:

```{r}
set.seed(123)
split_strat  <- initial_split(meats, prop = 0.8, 
                              strata = "water")
meats_train  <- training(split_strat)
meats_test   <- testing(split_strat)
```

### Training and fitting models using RMSE metric:

```{r}
my_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
my_metric <- "RMSE"

set.seed(123)
mod_lm <- train(water ~ .,
                     data = meats_train,
                     method = 'lm',
                     metric = my_metric,
                     preProcess = c("center", "scale"),
                     trControl = my_ctrl)

set.seed(123)
mod_enet <- train(water ~ .,
                      data = meats_train,
                      method = 'glmnet',
                      metric = my_metric,
                      preProcess = c("center", "scale"),
                      trControl = my_ctrl)

set.seed(123)
mod_nnet <- train(water ~ .,
                            data = meats_train,
                            method = "nnet",
                            preProcess = c("center", "scale"),
                            metric = my_metric,
                            trControl = my_ctrl,
                            trace = FALSE,
                            linout = TRUE)

set.seed(123)
mod_xgboost <- train(water ~ .,
                     data = meats_train,
                     method = 'xgbTree',
                     metric = my_metric,
                     trControl = my_ctrl)

set.seed(123)
mod_rf <- train(water ~ .,
                     data = meats_train,
                     method = 'rf',
                     metric = my_metric,
                     preProcess = c("center", "scale"),
                     trControl = my_ctrl)
```

```{r}
set.seed(123)
mod_pcr <- train(water ~ .,
                 data = meats_train,
                 method = "pcr",
                 trControl = my_ctrl,
                 preProcess = c("zv","center", "scale"),
                 #metric = my_metric,
                 tuneLength = 100)
```

### Comparision of models on resampled dataset:

```{r}
results = resamples(list(lm = mod_lm,
                         enet = mod_enet,
                         nnet = mod_nnet,
                         xgboost = mod_xgboost,
                         rf = mod_rf
                         ))
dotplot(results, metric = "RMSE")
```

`mod_nnet` has the lowest RMSE, by far, on the resampled dataset. 

### Performance on hold-out test data 

```{r}
test_data <- subset (meats_test, select = -c(water))
```

```{r}
pred_water_lm <- as.data.frame(predict(mod_lm, newdata = test_data))

compare_water_pred_lm <- cbind(meats_test, pred_water_lm)
names(compare_water_pred_lm)[ncol(compare_water_pred_lm)] <- "predicted_water_content"

compare_water_pred_lm %>% ggplot(aes(x = water, y = predicted_water_content))+
  geom_point()+
  geom_abline(slope=1, intercept=0, colour = "red", size = 0.8)+
  xlab("Observed water content")
###

pred_water_enet <- as.data.frame(predict(mod_enet, newdata = test_data))

compare_water_pred_enet <- cbind(meats_test, pred_water_enet)
names(compare_water_pred_enet)[ncol(compare_water_pred_enet)] <- "predicted_water_content"

compare_water_pred_enet %>% ggplot(aes(x = water, y = predicted_water_content))+
  geom_point()+
  geom_abline(slope=1, intercept=0, colour = "red", size = 0.8)+
  xlab("Observed water content")
###

pred_water_nnet <- as.data.frame(predict(mod_nnet, newdata = test_data))

compare_water_pred_nnet <- cbind(meats_test, pred_water_nnet)
names(compare_water_pred_nnet)[ncol(compare_water_pred_nnet)] <- "predicted_water_content"

compare_water_pred_nnet %>% ggplot(aes(x = water, y = predicted_water_content))+
  geom_point()+
  geom_abline(slope=1, intercept=0, colour = "red", size = 0.8)+
  xlab("Observed water content")
###

pred_water_xgb <- as.data.frame(predict(mod_xgboost, newdata = test_data))

compare_water_pred_xgb <- cbind(meats_test, pred_water_xgb)
names(compare_water_pred_xgb)[ncol(compare_water_pred_xgb)] <- "predicted_water_content"

compare_water_pred_xgb %>% ggplot(aes(x = water, y = predicted_water_content))+
  geom_point()+
  geom_abline(slope=1, intercept=0, colour = "red", size = 0.8)+
  xlab("Observed water content")
###

pred_water_rf <- as.data.frame(predict(mod_rf, newdata = test_data))

compare_water_pred_rf <- cbind(meats_test, pred_water_rf)
names(compare_water_pred_rf)[ncol(compare_water_pred_rf)] <- "predicted_water_content"

compare_water_pred_rf %>% ggplot(aes(x = water, y = predicted_water_content))+
  geom_point()+
  geom_abline(slope=1, intercept=0, colour = "red", size = 0.8)+
  xlab("Observed water content")
```

The neural network performs best on the test data too.

### Quantifying the model fits on test data set with RMSE

```{r}
#contour_data <- as.data.frame(matrix())
rmse_lm = rmse(actual = compare_water_pred_lm$water, predicted = compare_water_pred_lm$predicted_water_content)

rmse_enet = rmse(actual = compare_water_pred_enet$water, predicted = compare_water_pred_enet$predicted_water_content)

rmse_nnet = rmse(actual = compare_water_pred_nnet$water, predicted = compare_water_pred_nnet$predicted_water_content)

rmse_xgb = rmse(actual = compare_water_pred_xgb$water, predicted = compare_water_pred_xgb$predicted_water_content)

rmse_rf = rmse(actual = compare_water_pred_rf$water, predicted = compare_water_pred_rf$predicted_water_content)

rmse_lm
rmse_enet
rmse_nnet
rmse_xgb
rmse_rf
```

### Visualizing RMSE values of fits on test data

```{r}
my_Frame <- tribble(
  ~model  ,     ~RMSE,    
   "lm"   ,        2.06892,
   "enet" ,        2.4097,
   "nnet" ,        1.33733,
   "xgb"  ,        6.586896,
   "rf"   ,       5.98447
)

my_Frame %>% ggplot(aes(x = model, y = RMSE))+
  geom_point(size = 4)
```

### Looking at important variables according to the different models:

```{r}
plot(varImp(mod_lm), top = 20)
plot(varImp(mod_enet), top = 20)
plot(varImp(mod_nnet), top = 20)
plot(varImp(mod_xgboost), top = 20)
plot(varImp(mod_rf), top = 20)
#plot(varImp(mod_pcr), top = 20)
```

`x_040`, `x_042`, `x_041` are considered prominent by all models except the neural network. The neural network thinks `x_100` is indispensable and that `x_098` is also important.

But since the neural network seems to perform best on both training and test data, I choose its preferred variables to be more important.

```{r}
#for(i in 1:6){
#  viz_data[2] = var_2[i]
#  new_predicted_data <- predict(mod_nnet, viz_data)
  
#  assign(paste("viz_data", i, sep = "_"), cbind(viz_data, new_predicted_data))
  #viz_data <- cbind(viz_data, new_predicted_data)

#  temp_plot = v[i] %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()
  
#  assign(paste("p",i, sep=""), temp)
  #viz_data <- subset(viz_data, select = -predicted_water_content)
#}
```

### Predicting water content using 2 most important variables.

```{r}
# Six data points on the "facet" variable
var_2 = seq(min(meats[98]), max(meats[98]), length.out=6)

# Creating 6 new data frames each with one data point for the facet variable.
viz_data <- select(meats, x_100, x_098)
j=3

for(i in 1:100){
  if((i != 100) && (i!= 98)){
    viz_data[j] = median(as.data.frame(meats)[,i])
    colnames(viz_data)[j] <- colnames(meats)[i]
    j = j+1
  }
}

# viz_data_1
## set value of facet variable
viz_data[2] = var_2[1]
## predict response using updated dataset
new_predicted_data <- predict(mod_nnet, viz_data)
viz_data_1 <- cbind(viz_data, new_predicted_data)
##renaming column of predictde water content
names(viz_data_1)[ncol(viz_data_1)] <- "predicted_water_content"
## Assigning the plot to a variable
p1 = viz_data_1 %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()

# viz_data_1
viz_data[2] = var_2[2]
new_predicted_data <- predict(mod_nnet, viz_data)
viz_data_2 <- cbind(viz_data, new_predicted_data)
names(viz_data_2)[ncol(viz_data_2)] <- "predicted_water_content"
p2 = viz_data_2 %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()

# viz_data_1
viz_data[2] = var_2[3]
new_predicted_data <- predict(mod_nnet, viz_data)
viz_data_3 <- cbind(viz_data, new_predicted_data)
names(viz_data_3)[ncol(viz_data_3)] <- "predicted_water_content"
p3 = viz_data_3 %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()

# viz_data_1
viz_data[2] = var_2[4]
new_predicted_data <- predict(mod_nnet, viz_data)
viz_data_4 <- cbind(viz_data, new_predicted_data)
names(viz_data_4)[ncol(viz_data_4)] <- "predicted_water_content"
p4 = viz_data_4 %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()

# viz_data_1
viz_data[2] = var_2[5]
new_predicted_data <- predict(mod_nnet, viz_data)
viz_data_5 <- cbind(viz_data, new_predicted_data)
names(viz_data_5)[ncol(viz_data_5)] <- "predicted_water_content"
p5 = viz_data_5 %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()

# viz_data_1
viz_data[2] = var_2[6]
new_predicted_data <- predict(mod_nnet, viz_data)
viz_data_6 <- cbind(viz_data, new_predicted_data)
names(viz_data_6)[ncol(viz_data_6)] <- "predicted_water_content"
p6 = viz_data_6 %>% ggplot(aes(x = x_100, y = predicted_water_content))+geom_point()
```

### Plotting water content wrt `x_100` as primary variable and different values of facet variable `x_098`:

```{r}
(p1+p2+p3) / (p4+p5+p6)
```